\textls[-15]{\thetitle \mbox{(Technical Report)}}

### 2022-07-03 coq-example.tex
Small fixes through out.

Our emphasis here is that, we singled out some language features from
STLC and prove the type safety for each feature separately, following
the style of the examples in MTC~\cite{delaware2013,forsta2020}.
And we use a \textit{not-yet-formally-defined} \textbf{mix-in} feature
to mix the semantics and properties of two language features---product
and boolean (in \cref{fig:STLC-example})---with vanilla STLC. This way,
we can say precisely, \textit{a programming language feature itself is a
piece of data/inheritance judgement/inherited family}. Compared with MTC,
our example uses small-step operational semantics\YZ{Does MTC not define a small-step operational semantics for STLC?}\EDJreply{I thought they use big step with fuel. Though Coq a la carte use small-step in one example} by exploiting
extensible inductive types.

Most of the proof and computation is directly adapted
from the one in Software Foundation, resulting in a similarly accessible proof, while the proof for each extended feature are scattered in the extended families resulting better modularity. One kind of examples is about the computation, and here we mainly focus on substitution functions. Substitutions dealing with product and projection are defined in the corresponding children family. This is mainly done by inheriting and extending the handler family from the parents. The other kind of examples are about proof-irrelevant propositions and we directly use "(Extend) FTheorem" to handle them. In these cases we don't even need to additionally handle a auxillary handler family like we did for substitution---we simply use "Extend FTheorem" and our plugin will prepare the goal to prove. At the end, it will look like scattered proof scripts in a modular manner.
For example, the proof on progress on product projection is carried out in the children family that extend "tm" with products. \YZ{Can you say more about how your plugin enables more code reuse and modularity?}\EDJreply{please check}
% In related work, Coq/Metatheory a la carte/Tion embedding can be emphasized as a more "semantic approach" because they encode the meaning using a special design pattern (for example, open recursive inductive type for extension), compared to our more syntactic approach. Their advantage is the transferability of this technique accross different proof assistants, and their disadvantage is that their approach are less accessible and unfriendly to amateur Coq users---which can be reflected from the distance of their approach and text-book Software Foundation proof.

% We have problem on inversion lemma. Check if it is the same problem as MTC
We use "Closing Fact" to state and prove the inversion lemma instead of
using the extensible proving mechanism "FTheorem". The reasons are that
(1) it introduces much less boilerplate code because the proof for these
inversion lemmas should be just simple case analysis and we should rely
on Coq to generate most of the boilerplate code;\YZ{I was under the impression that the plugin could not auto-generate Closing Facts or their proofs, no?}\EDJreply{I have make the above Closing Fact section more about this detail. Please check.} (2) it shouldn't bother us in the future
because any extension on the syntax should still satisfy these inversion
lemmas; (3) most importantly, we believe this inversion ``lemma'' should
be part of the definition of the syntax instead of considering the
syntax as a mere concrete inductive type. We should postulate this
inversion ``lemma'' like \textit{a constraint} and post-hoc-ly verify
that our inductive definition does satisfy the constraint, which is
exactly what we expect from "Closing Fact".

@@

Another difference comes from the fact that we need to use "Closing
Fact" to manually verify the computational rules for each recursor and
use tactics to ``run'' the recursor by repeatedly "rewrite" using those
verified computational rules. This part is possible to be automatically
handled by the plugin, but still it will require propositional
computational rules.

\subsection{Abstract Interpreters for \texttt{Imp}}
The second example is adapted and modified from \citet{zm2017}.
Contrary to our first example, here, we use a big-step interpreter with fuel
as the operational semantics of an imperative language with
side effects, and we specify the abstract interpretation and prove its
soundness, with some postulation on both computation and property.
Then we extend the base language with new features, and we instantiate
the postulation on computation for both concrete interpreters and
abstract interpreters.
Thanks to the compilation to Coq, we can directly run the resulting
abstract interpreters.

@@

\begin{figure}
  \includegraphics[width=\columnwidth]{coqexmaple/Family-Lang-Imp2.pdf}
  \caption{Inheritance Diagram for Abstract Interpretation Example. Black arrows are inheritance. Purple arrow are extending. Green arrows are overriding. Grey box are sealed families.}\label{fig:abstract-interpretation-example}
\end{figure}

>>>>

Our emphasis here is that, we singled out some language features from
STLC and prove the type safety for each feature separately, following
the style of the examples in MTC~\cite{delaware2013,forsta2020}.
And we use a \textit{not-yet-formally-defined} \textbf{mix-in} feature
to mix the semantics and properties of two language features---product
and boolean (in \cref{fig:STLC-example})---with vanilla STLC.

This \textit{mixin} is implemented roughly by making inheritance judgement/data supporting \textit{weakening} and composition. Then, using notation from metatheory, given $\goodInh{\cdot}{i_1}{\sigma_1}{\sigma_2}$ and $\goodInh{\cdot}{i_2}{\sigma_1}{\sigma_3}$, we can weaken $i_2$ to $\goodInh{\cdot}{i_2'}{\sigma_2}{\sigma_3'}$ and the composition of $i_1$ and $i_2'$ is the resulting mixin---this implementation breaks commutativity and will need to solve (design) some compatibility solution: for example, when $i_1$ and $i_2$ are both extending with field of the same name, do we signal error or use $i_2$'s extension as considering it as an `override'? However, our current example of STLC only uses a subset of the mixin that is with a clear semantic (i.e. every sensible design of mixin should agree on).

Using this example,
we hope one day we can say precisely, \textit{a programming language feature itself is a
piece of data/inheritance judgement/inherited family}. Compared with MTC,
our example uses small-step operational semantics\YZ{Does MTC not define a small-step operational semantics for STLC?}\EDJreply{I thought they use big step with fuel. Though Coq a la carte use small-step in one example} by exploiting
extensible inductive types.

Most of the proof and computation are directly adapted
from the one in Software Foundation, resulting in a similarly accessible proof, while the proof for each extended feature are scattered in the extended families resulting better modularity. One kind of examples is about the computation, for example, substitution functions. Substitutions dealing with product and projection are defined in the corresponding children family. This is mainly done by inheriting and extending the handler family from the parents. The other kind of examples are about proof-irrelevant propositions and we directly use "(Extend) FTheorem" to handle them. In these cases we don't need to additionally define a auxillary handler family like we did for substitution---we simply use "Extend FTheorem" and our plugin will prepare the goal to prove. At the end, it will look like scattered proof scripts in a modular manner.
For example, the proof on progress on product projection is carried out in the children family that extend "tm" with products. \YZ{Can you say more about how your plugin enables more code reuse and modularity?}\EDJreply{please check}
% In related work, Coq/Metatheory a la carte/Tion embedding can be emphasized as a more "semantic approach" because they encode the meaning using a special design pattern (for example, open recursive inductive type for extension), compared to our more syntactic approach. Their advantage is the transferability of this technique accross different proof assistants, and their disadvantage is that their approach are less accessible and unfriendly to amateur Coq users---which can be reflected from the distance of their approach and text-book Software Foundation proof.

% We have problem on inversion lemma. Check if it is the same problem as MTC
We use "Closing Fact" to state and prove the inversion lemma instead of
using the extensible proving mechanism "FTheorem". The reasons are that
(1) it introduces much less boilerplate code because the proof for these
inversion lemmas should be just simple case analysis and we should rely
on Coq to generate most of the boilerplate code;\YZ{I was under the impression that the plugin could not auto-generate Closing Facts or their proofs, no?}\EDJreply{I have make the above Closing Fact section more about this detail. Please check.} (2) it shouldn't bother us in the future
because any extension on the syntax should still satisfy these inversion
lemmas; (3) most importantly, we believe this inversion ``lemma'' should
be part of the definition of the syntax instead of considering the
syntax as a mere concrete inductive type. We should postulate this
inversion ``lemma'' like \textit{a constraint} and post-hoc-ly verify
that our inductive definition does satisfy the constraint, which is
exactly what we expect from "Closing Fact"---if directly working on inductive definition doesn't bring us benefit, then decompose the property (the constraint) with the concrete definition might be a good idea.

@@

Another difference comes from the fact that we need to use "Closing
Fact" to manually verify the computational rules for each recursor and
use tactics to ``run'' the recursor by repeatedly "rewrite" using those
verified computational rules. This part is possible to be automatically
generated by the plugin, but still it will require propositional
computational rules.

\subsection{Abstract Interpreters for \texttt{Imp}}
The second example is adapted and modified from \citet{zm2017}.
Contrary to our first example, here, we use a big-step interpreter with fuel
as the operational semantics of an imperative language with
side effects, and we specify the abstract interpretation and prove its
soundness, with some postulation on both computation and property.
Then we extend the base language with new features, and we instantiate
the postulation on computation for
abstract interpreters.
Thanks to the compilation to Coq module, we can directly run the resulting
abstract interpreters.

\begin{figure}[!htb]
  \includegraphics[width=\columnwidth]{coqexmaple/Family-Lang-Imp2.pdf}
  \caption{Inheritance Diagram for Abstract Interpretation Example. Black arrows are inheritance. Purple arrow are extending. Green arrows are overriding. Grey box are sealed families.}\label{fig:abstract-interpretation-example}
\end{figure}

@@

### 2022-07-03 related-work.tex
Small fixes through out.

\textbf{Modular Mechanized Metatheory: Syntactic Approach} We group the work that achieves modular mechanized metatheory by extending surface syntax of proof assistants as \textit{syntactic approach}.

The closest work to ours might be \citet{boite2004proo}--they try to support \textit{extended} inductive type and related proof reuse in Coq. This biggest difference lies in our (incomplete) adaption of Family Polymorphism -- apart from supporting proof reuse and extension by inheritance, we also support implementation swapping using overriding, and we also argue that Family Polymorphism can provide a better and cleaner way of organizing mechanized metatheory development. The other difference lies in the handling of the extension of the inductive type. In our paradigm, the discussion of \textit{extensible} inductive type and \textit{partial recursor} is unavoidable, which characterizes all possible extension of a given inductive type. In their work, without using Family to organized the code, the extended inductive type can be considered as a different inductive type than the original one, possibly leading to more expression problem~\cite{wadler-ep}. 
% Another similar work is proof weaving \citep{mulhern2006proof} but I don't see a evaluation section in that paper

The other work

\textbf{Modular Mechanized Metatheory: Semantic Approach} We group the work using certain algebraic encoding (e.g. Church Encoding, datatype-generic programming) without modify the surface syntax of the proof assistants as \textit{semantic approach}, that includes \citet{delaware2013,forsta2020,liwei2022,schwaab2013modular, keuchel2013generic} and so on. Their advantages are various: this technique is more general and thus can be applied across different (dependent-typed) proof assistants and this technique can be easily distributed like a library and thus doesn't need to alter the core of the proof assistants. The downsides are mainly about the threshold: they are not friendly for the junior Coq development and these indirect encoding of datatype might lead to worse readability and accessibility compared to the direct style like ours.

>>>>

\textbf{Modular Mechanized Metatheory: Linguistic Approach} We group the work that achieves modular mechanized metatheory by extending surface syntax of proof assistants to provide better abstraction as \textit{linguistic approach}.

The closest work to ours might be \citet{boite2004proof}--they try to support \textit{extended} inductive type and related proof reuse in Coq. This biggest difference lies in our (incomplete) adaption of Family Polymorphism -- apart from supporting proof reuse and extension by inheritance, we also support implementation swapping using overriding, and we also argue that Family Polymorphism can provide a better and cleaner way of organizing mechanized metatheory development a well as solving related expression problem. The other difference lies in the handling of the extension of the inductive type. In our paradigm, the discussion of \textit{extensible} inductive type and \textit{partial recursor} is unavoidable, which characterizes all possible extension of a given inductive type. In their work, without using Family to organized the code, the extended inductive type can be considered as a different inductive type than the original one, possibly leading to more expression problem~\cite{wadler-ep}. The situation is similar for \citet{levin2003tinkertype}---they aim for modular manipulation of metatheory but didn't consider expression problem.  What's more, our implementation is based on Coq plugin and is thus potentially more friendly to Coq community.
% Another similar work is proof weaving \citep{mulhern2006proof} but I don't see a evaluation section in that paper

\textbf{Modular Mechanized Metatheory: Semantic Approach} We group the work using the semantic as encoding (e.g. Church Encoding, datatype-generic programming) to achieve certain design patterns without modify the surface syntax of the proof assistants as \textit{semantic approach}, that includes \citet{delaware2013,forsta2020,liwei2022,schwaab2013modular, keuchel2013generic} and so on. Their advantages are various: their techniques are more general and thus can be applied across different (dependent-typed) proof assistants and their technique can be easily distributed like a library and thus doesn't need to alter the core of the proof assistants. The downsides are mainly about the threshold: they are not friendly for the junior Coq development and these indirect encoding of datatype might lead to worse readability and accessibility compared to the direct style like ours.
